Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-01 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: polishing text

\documentclass[twocolumn,prd,aps,superscriptaddress,preprintnumbers,tightenlines,showpacs,nofootinbib,eqsecnum,amsfonts,amsmath,longbibliography]{revtex4-1}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
\usepackage{txfonts}
% \usepackage{newtxtext}
% \usepackage{newtxmath}
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
\usepackage{mathrsfs}
\usepackage{makecell}
\usepackage{microtype}

@@

\newcommand{\h}{\mathpzc{h}}
\newcommand{\Hhat}{\hat{\mathpzc{H}}}
\newcommand{\B}{\mathpzc{B}}
\newcommand{\hlm}{\mathpzc{h}_{\ell m}}
\newcommand{\xilm}{\xi_{\ell m}}
\newcommand{\Ylm}{{Y}^{-2}_{\ell m}}
\newcommand{\Y}{{Y}^{-2}}
\newcommand{\hc}{h_\times}
\newcommand{\hp}{h_+}
\newcommand{\Fc}{F_\times}
\newcommand{\Fp}{F_+}
\newcommand{\Mf}{M_f}
\newcommand{\cA}{\mathpzc{A}}
\newcommand{\lm}{_{\ell m}}
\newcommand{\deff}{d_\mathrm{eff}}
\newcommand{\rmi}{\mathrm{i}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bxi}{\bm{\xi}}
\newcommand{\bxigr}{\bm{\xi}_{\text{GR}}}
\newcommand{\bxingr}{\bm{\xi}_{\text{nGR}}}
\newcommand{\bzeta}{\bm{\zeta}}
\newcommand{\bs}[1]{\bm{\vec{S}_{#1}}}
\newcommand{\Mo}{M_{\odot}}
\newcommand{\FFe}{\mathrm{FF}_\mathrm{eff}}
\newcommand{\FF}{\mathrm{FF}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\rhoopt}{\rho_\mathrm{opt}}
\newcommand{\rhosubopt}{\rho_\mathrm{subopt}}
\newcommand{\fqnm}{f}
\newcommand{\sigmaqnm}{\sigma}
\newcommand{\n}{\mathbf{n}}
\newcommand*{\skymapscale}{0.5}
\newcommand*{\paramestscale}{0.455}
\newcommand{\df}[1]{\delta f_{\text{#1}}}
\newcommand{\dtau}[1]{\delta \tau_{\text{#1}}}
\newcommand{\fngr}[1]{f_{\text{#1}}}
\newcommand{\taungr}[1]{\tau_{\text{#1}}}
\newcommand{\fgr}[1]{f ^{\text{GR}}_{\text{#1}}}
\newcommand{\taugr}[1]{\tau ^{\text{GR}}_{\text{#1}}}
\newcommand{\pSEOB}{\texttt{pSEOBNR}}
\newcommand{\SEOB}{\texttt{SEOBNR}}
\newcommand{\gm}{\mathfrak{m}}

@@

The polarisations of the GW signal (in the observer's frame) are,
%
\begin{equation}
h_+(\iota,\varphi_0;t ) - i h_\times(\iota,\varphi_0;t) = \sum_{l, m} {}_{-\!2}Y_{l m}(\iota,\varphi_0)\, h_{l m}(t)\,,
\end{equation}
%
where ${}_{-\!2}Y_{l m}(\iota,\varphi_0)$ are the $-2$ spin-weighted spherical harmonics described by standard angular dependence indices $(l,m)$.\footnote{Note that we use $l$ to denote the angular dependence index and $\ell$ to denote the coupling constant in Sec.~\ref{sec:review_parspec}.} As our baseline model, we use the computationally efficient (time-domain) multipolar waveform model for quasicircular spin-aligned BBH mergers described in \cite{Mihaylov:2021bpf} (henceforth referred to as \SEOB{} \footnote{This waveform model is available in \texttt{LALSuite} \cite{lalsuite} as the \texttt{SEOBNRv4HM\_PA} waveform approximant.}) which contains the modes, $(l, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, and $(5,5)$~\cite{Cotesta:2018fcv,Mihaylov:2021bpf}. The model uses an effective-one-body approach combined with a post-adiabatic solution of the equations of motion to describe the inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$. An accurate description of the merger is incorporated through calibration with NR simulations, as described in~\cite{Cotesta:2018fcv}, along with information for the merger and ringdown phases, from BH perturbation theory. The merger-ringdown waveform, $h_{l m}^\mathrm{merger-RD}$, is then stitched to inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$ at a certain time $t = t^{\textrm{match}}_{l m}$, as
%
\begin{equation}
h_{l m}(t) = h_{l m}^\mathrm{insp-plunge}\, \theta(t_\mathrm{match}^{l m} - t) + h_{l m}^\mathrm{merger-RD}\,\theta(t-t_\mathrm{match}^{l m})\,,
\end{equation}
%
where $\theta(t)$ is the Heaviside step function. The merger-ringdown waveform is expressed as an exponentially damped sinusoid ~\citep{Bohe:2016gbl,Cotesta:2018fcv,Mihaylov:2021bpf}
%
\begin{equation}
\label{RD}
h_{l m}^{\textrm{merger-RD}}(t) = \nu \ \tilde{A}_{l m}(t)\ e^{i \tilde{\phi}_{l m}(t)} \ e^{-i \sigma_{l m 0}(t-t^{\textrm{match}}_{l m})},
\end{equation}
%
where
%
\begin{align}
\sigma_{l m 0} &= {\rm Re}(\sigma_{l m 0}) + i \, {\rm Im}(\sigma_{l m 0})\,,
\nonumber \\
&= \omega_{l m 0} - i / \tau_{l m 0}\,,
\end{align}
%
are the complex frequencies of the fundamental (n=0 overtone) QNMs of the remnant BH. The frequencies $\omega_{l m 0}$ and damping times $\tau_{l m 0}$ of the $(l, m)$ QNM can be read off from the real and imaginary parts of  $\sigma_{l m0}$ respetively. The functions $\tilde{A}_{l m}(t)$ and $\tilde{\phi}_{l m}(t)$ are defined in~\cite{Bohe:2016gbl,Cotesta:2018fcv}.

@@

The \pSEOB{} model, as described in Sec.~\ref{sec:review_pSEOB}, is an \emph{IMR} model which infers the properties of the underlying GW signal, including (independently) its ringdown properties, using the Bayesian formalism above. Naturally, the most promising candidates for our analyses are high-mass \emph{and} loud GW observations with a significant signal-to-noise ratio (SNR) in the post-merger part of the signal. The latest LVK GW catalog~\cite{LIGOScientific:2021djp} reported 90 observed signals not all of which are relevant for a BH ringdown analysis. In fact, in an accompanying paper~\cite{LIGOScientific:2021sio}, the \textsc{pSEOBNRv4HM} analysis~\footnote{Sec.VIII A.2 in ~\cite{LIGOScientific:2021sio}}, which is most similar to the \pSEOB{} model presented in this paper, identified two events which provided the strongest bounds on measurements of the dominant $(220)$ QNM -- GW150914~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp}. These two events, with a total mass of $XX \Mo$ and $XX \Mo$ respectively, are extremely similar in their source properties. These are also two of the loudest BBH signals observed to date with a total network SNR of XX and XX respectively. Moreover, and what is more relevant for our analysis, is their post-inspiral (merger-ringdown) SNRs, which at XX and XX respectively (see the columns for $\rho_{\text{post-insp}}$ in Table III of~\cite{LIGOScientific:2019fpa} and Table IV of ~\cite{LIGOScientific:2021sio}). In this paper, we are going to focus on these two GW events as our probes of the BH ringdown in alternative theories of gravity.

@@

First, as we explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be considered as an effective field theory, meaning it should be
considered valid only below an energy scale, or equivalently, a length scale.
%
As a cut-off length scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon; \, \mathfrak{m}) = \varepsilon \,  (G \mathfrak{m} / c^2)\,,
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the masses involved in the problem.
%
We also momentarily restored factors of $c$ and $G$ to emphasize that $\Lambda_{\rm EFT}$ has
dimensions of length and hence can be compared to the coupling strength $\ell$.
%
The parameter $\Lambda_{\rm EFT}$ was used
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} with $\varepsilon = 1/2$,
but here we  $\varepsilon \in [0, 1]$.
%
We will say that a bound has been placed on $\ell$, if most of the support of
the posterior distribution function $P(\ell)$ is in the interval $[0, \, \Lambda_{\rm EFT}(\varepsilon; \, \mathfrak{m})]$.
%
In practice, this can be quantified by the cumulative distribution function
(CDF) associated to the marginalized posterior distribution $P(\ell)$,
%
\begin{equation}
P(\ell \leqslant \ell') = \int_{-\infty}^{\,\ell'} P(\ell') \, \dd \ell'\,.
\label{eq:def_cdf}
\end{equation}
%
For instance, we will demand that for a bound with 90\% credibility to be meaningfully placed that,
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell \leqslant \Lambda_{\rm EFT}) \leqslant 0.9 \,.
\label{eq:eft_bound}
\end{equation}

%
Another question to consider is the following: what is the mass $\mathfrak{m}$ that we should use?
%
In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, which attempted to
constrain dCS and shift-symmetric sGB theories with the \emph{inspiral} part of
signal alone, it was natural to chose the secondary's mass $m_2$ as the most
conservative choice, for it is by definition the smallest mass and hence
places the lowest threshold for the validity of either theory as an EFT.
%
In our problem, the answer is not as clear. On the one hand, since we are
interested in the ringdown part of the signal, it is natural to use the
final mass $M_f$ to compute the cut-off $\Lambda_{\rm EFT}$.
%
On the other hand, one may argue that the modified gravity theory under study
should be able to predict a full inspiral, merger, and ringdown of the black
hole binary, and thus the same most conservative $m_2$ should be used.
%
Here we adopt an agnostic view and consider both masses, $m_2$ and $M_f$, to
determine $\Lambda_{\rm EFT}$. We will then compare how different assumptions
yield to different interpretations of the results of our parameter estimations.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The cumulative distribution function evaluated at the
threshold $T(\varepsilon;\, m)$ EdGB gravity as a function of the parameter
$\varepsilon$ for both (blue curves) $m = m_2$, the secondary's source
mass, and (orange curves) $m = M_{f}$, the remnant's source mass, without
(solid curve) and with (dashed curves) linear in spin QNM corrections. The
horizontal lines lay at 0.68 and 0.90.
}
\label{fig:dcs_cdf}
\end{figure}

@@

In Fig.~\ref{fig:dcs_cdf} we show the cumulative distribution function (CDF) as
evaluates with Eq.~\eqref{eq:def_cdf}, with $\ell' = \Lambda_{\rm EFT}$,
for both GW150914 (top panel) and GW200129 (bottom panel), where $\Lambda_{\rm EFT}$
is calculated with either $\mathfrak{m} = m_2$ (blue curves) or $M_{f}$ (orange curves),
and with $\varepsilon \in [0,\, 1]$.
%
For GW150914, we see that with $\gm = m_2$ that Eq.~\eqref{eq:eft_bound} is not
satisfied unless $\varepsilon \approx 0.9$ (with only $j=0$ corrections) and
$\varepsilon \approx 0.7$ (with both $j=0$ and $1$ corrections).
%
The situation is different if we use $\gm = M_f$. In this case, we find that
with or without spin corrections that Eq.~\eqref{eq:eft_bound} can be satisfied
with $\varepsilon \leq 1/2$, i.e., below the criteria used
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}.
%
This means that with our model's assumptions and using the remnant's source mass $M_f$ to set the
cut-off scale that we can claim an upper bound
%
\begin{equation}
\ell_{\rm CS} \leqslant \Lambda_{\rm EFT}(1/2; \, M_{f}) = \dots \,.
\quad \textrm{at 90\% credibility},
\end{equation}
%
on dCS gravity and would constitute the first bound on this theory with
gravitational-wave observations alone.
%
These conclusions for GW150914 are shared with GW200129.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/dcs_cdf_varying_threshold.pdf}
\caption{(Color online).~The cumulative distribution function evaluated at the threshold $T(\varepsilon;\, m)$
as a function of the parameter $\varepsilon$ for both (blue curves) $m = m_2$, the secondary's source mass,
and (orange curves) $m = M_{f}$, the remnant's source mass, without (solid curve) and with (dashed curves)
linear in spin QNM corrections. The horizontal lines lay at 0.68 and 0.90.
}
\label{fig:dcs_cdf}
\end{figure}

>>>>

\documentclass[twocolumn,prd,aps,superscriptaddress,preprintnumbers,tightenlines,showpacs,nofootinbib,amsfonts,amsmath,longbibliography]{revtex4-1}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
\usepackage{txfonts}
% \usepackage{newtxtext}
% \usepackage{newtxmath}
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
\usepackage{mathrsfs}
\usepackage{makecell}
\usepackage{microtype}

@@

\newcommand{\h}{\mathpzc{h}}
\newcommand{\Hhat}{\hat{\mathpzc{H}}}
\newcommand{\B}{\mathpzc{B}}
\newcommand{\hlm}{\mathpzc{h}_{\ell m}}
\newcommand{\xilm}{\xi_{\ell m}}
\newcommand{\Ylm}{{Y}^{-2}_{\ell m}}
\newcommand{\Y}{{Y}^{-2}}
\newcommand{\hc}{h_\times}
\newcommand{\hp}{h_+}
\newcommand{\Fc}{F_\times}
\newcommand{\Fp}{F_+}
\newcommand{\Mf}{M_f}
\newcommand{\cA}{\mathpzc{A}}
\newcommand{\lm}{_{\ell m}}
\newcommand{\deff}{d_\mathrm{eff}}
\newcommand{\rmi}{\mathrm{i}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bxi}{\bm{\xi}}
\newcommand{\bxigr}{\bm{\xi}_{\text{GR}}}
\newcommand{\bxingr}{\bm{\xi}_{\text{nGR}}}
\newcommand{\bzeta}{\bm{\zeta}}
\newcommand{\bs}[1]{\bm{\vec{S}_{#1}}}
\newcommand{\Mo}{M_{\odot}}
\newcommand{\FFe}{\mathrm{FF}_\mathrm{eff}}
\newcommand{\FF}{\mathrm{FF}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\rhoopt}{\rho_\mathrm{opt}}
\newcommand{\rhosubopt}{\rho_\mathrm{subopt}}
\newcommand{\fqnm}{f}
\newcommand{\sigmaqnm}{\sigma}
\newcommand{\n}{\mathbf{n}}
\newcommand*{\skymapscale}{0.5}
\newcommand*{\paramestscale}{0.455}
\newcommand{\df}[1]{\delta f_{\text{#1}}}
\newcommand{\dtau}[1]{\delta \tau_{\text{#1}}}
\newcommand{\fngr}[1]{f_{\text{#1}}}
\newcommand{\taungr}[1]{\tau_{\text{#1}}}
\newcommand{\fgr}[1]{f ^{\text{GR}}_{\text{#1}}}
\newcommand{\taugr}[1]{\tau ^{\text{GR}}_{\text{#1}}}
\newcommand{\pSEOB}{\texttt{pSEOBNR}}
\newcommand{\SEOB}{\texttt{SEOBNR}}
\newcommand{\gm}{\mathfrak{m}}
\newcommand{\msun}{~{\rm M}_{\odot}}

@@

The polarisations of the GW signal (in the observer's frame) are,
%
\begin{equation}
h_+(\iota,\varphi_0;t ) - i h_\times(\iota,\varphi_0;t) = \sum_{l, m} {}_{-\!2}Y_{l m}(\iota,\varphi_0)\, h_{l m}(t)\,,
\end{equation}
%
where ${}_{-\!2}Y_{l m}(\iota,\varphi_0)$ are the $-2$ spin-weighted spherical harmonics described by standard angular dependence indices $(l,m)$.\footnote{Note that we use $l$ to denote the angular dependence index and $\ell$ to denote the coupling constant in Sec.~\ref{sec:review_parspec}.} As our baseline model, we use the computationally efficient (time-domain) multipolar waveform model for quasicircular spin-aligned BBH mergers described in \cite{Mihaylov:2021bpf} (henceforth referred to as \SEOB{}\footnote{This waveform model is available in \texttt{LALSuite} \cite{lalsuite} as the \texttt{SEOBNRv4HM\_PA} waveform approximant.}) which contains the modes, $(l, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, and $(5,5)$~\cite{Cotesta:2018fcv,Mihaylov:2021bpf}. The model uses an effective-one-body approach combined with a post-adiabatic solution of the equations of motion to describe the inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$. An accurate description of the merger is incorporated through calibration with NR simulations, as described in~\cite{Cotesta:2018fcv}, along with information for the merger and ringdown phases, from BH perturbation theory. The merger-ringdown waveform, $h_{l m}^\mathrm{merger-RD}$, is then stitched to inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$ at a certain time $t = t^{\textrm{match}}_{l m}$, as
%
\begin{equation}
h_{l m}(t) = h_{l m}^\mathrm{insp-plunge}\, \theta(t_\mathrm{match}^{l m} - t) + h_{l m}^\mathrm{merger-RD}\,\theta(t-t_\mathrm{match}^{l m})\,,
\end{equation}
%
where $\theta(t)$ is the Heaviside step function. The merger-ringdown waveform is expressed as an exponentially damped sinusoid ~\citep{Bohe:2016gbl,Cotesta:2018fcv,Mihaylov:2021bpf}
%
\begin{equation}
\label{RD}
h_{l m}^{\textrm{merger-RD}}(t) = \nu \ \tilde{A}_{l m}(t)\ e^{i \tilde{\phi}_{l m}(t)} \ e^{-i \sigma_{l m 0}(t-t^{\textrm{match}}_{l m})},
\end{equation}
%
where
%
\begin{align}
\sigma_{l m 0} &= {\rm Re}(\sigma_{l m 0}) + i \, {\rm Im}(\sigma_{l m 0})\,,
\nonumber \\
&= \omega_{l m 0} - i / \tau_{l m 0}\,,
\end{align}
%
are the complex frequencies of the fundamental (n=0 overtone) QNMs of the remnant BH. The frequencies $\omega_{l m 0}$ and damping times $\tau_{l m 0}$ of the $(l, m)$ QNM can be read off from the real and imaginary parts of  $\sigma_{l m0}$ respetively. The functions $\tilde{A}_{l m}(t)$ and $\tilde{\phi}_{l m}(t)$ are defined in~\cite{Bohe:2016gbl,Cotesta:2018fcv}.

@@

The \pSEOB{} model, as described in Sec.~\ref{sec:review_pSEOB}, is an \emph{IMR} model which infers the properties of the underlying GW signal, including (independently) its ringdown properties, using the Bayesian formalism above. Naturally, the most promising candidates for our analyses are high-mass \emph{and} loud GW observations with a significant signal-to-noise ratio (SNR) in the post-merger part of the signal. The latest LVK GW catalog~\cite{LIGOScientific:2021djp} reported 90 observed signals not all of which are relevant for a BH ringdown analysis. In fact, in an accompanying paper~\cite{LIGOScientific:2021sio}, the \textsc{pSEOBNRv4HM} analysis~\footnote{Sec.VIII A.2 in ~\cite{LIGOScientific:2021sio}}, which is most similar to the \pSEOB{} model presented in this paper, identified two events which provided the strongest bounds on measurements of the dominant $(220)$ QNM -- GW150914~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp}. These two events, with a total mass of $XX \Mo$ and $XX \Mo$ respectively, are extremely similar in their source properties. These are also two of the loudest BBH signals observed to date with a total network SNR of XX and XX respectively. Moreover, and what is more relevant for our analysis, is their post-inspiral (merger-ringdown) SNRs, which at XX and XX respectively (see the columns for $\rho_{\text{post-insp}}$ in Table III of~\cite{LIGOScientific:2019fpa} and Table IV of ~\cite{LIGOScientific:2021sio}). In this paper, we are going to focus on these two GW events as our probes of the BH ringdown in modified theories of gravity.

@@

First, as we explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be considered as an effective field theory, meaning it should be
considered valid only below an energy scale, or equivalently, a length scale.
%
As a cut-off length scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, G \mathfrak{m} / c^2 \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the masses involved in the problem.
%
We also momentarily restored factors of $c$ and $G$ to emphasize that $\Lambda_{\rm EFT}$ has
dimensions of length and hence can be compared to the coupling strength $\ell$.
%
In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
but here we consider $\varepsilon \in [0, 1]$.
%
We will say that a bound has been placed on $\ell$, if most of the support of
the posterior distribution function $P(\ell)$ is in the interval
$[0, \, \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated to the marginalized posterior distribution $P(\ell)$,
%
\begin{equation}
P(\ell \leqslant \ell') = \int_{-\infty}^{\,\ell'} P(\ell') \, \dd \ell'\,.
\label{eq:def_cdf}
\end{equation}
%
For instance, we will demand that for a bound with 90\% credibility to be meaningfully placed that,
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell \leqslant \Lambda_{\rm EFT}) \leqslant 0.9 \,,
\label{eq:eft_bound}
\end{equation}
%
and likewise for other percentiles.

Another question we must consider is the following: what is the mass $\mathfrak{m}$ that we should use
in Eq.~\eqref{eq:def_cutoff}?
%
In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, which attempted to
constrain dCS and shift-symmetric sGB theories with the \emph{inspiral} part of
the GW signal alone, it was natural to chose the secondary's mass $m_2$ as the most
conservative choice, for it is by definition the smallest mass and hence
places the lowest threshold for the validity of either theory as an EFT.
%
In our problem, the answer is not as clear. On the one hand, since we are
interested in the ringdown part of the signal, it is natural to use the
final mass $M_f$ to compute the cut-off $\Lambda_{\rm EFT}$.
%
On the other hand, one may argue that the modified gravity theory under study
should be able to predict a full inspiral, merger, and ringdown of the black
hole binary before we can even make such a test, and thus the same, more conservative choice
$\gm = m_2$ should be used.
%
Here we adopt an agnostic view to this question and consider \emph{both} masses, $m_2$ and $M_f$, to
determine $\Lambda_{\rm EFT}$. We will then compare how different assumptions
yield to different interpretations of the results of our parameter estimations.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The cumulative distribution function evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon,\gm)$ EdGB gravity as a function of the parameter
$\varepsilon$ for both (blue curves) $m = m_2$, the secondary's source
mass, and (orange curves) $m = M_{f}$, the remnant's source mass, without
(solid curve) and with (dashed curves) linear in spin QNM corrections. The
horizontal lines lay at 0.68 and 0.90.
}
\label{fig:dcs_cdf}
\end{figure}

@@

In Fig.~\ref{fig:dcs_cdf} we show the cumulative distribution function (CDF) as
evaluates with Eq.~\eqref{eq:def_cdf}, with $\ell' = \Lambda_{\rm EFT}$,
for both GW150914 (top panel) and GW200129 (bottom panel), where $\Lambda_{\rm EFT}$
is calculated with either $\mathfrak{m} = m_2$ (blue curves) or $M_{f}$ (orange curves),
and with $\varepsilon \in [0,\, 1]$.
%
For GW150914, we see that with $\gm = m_2$ that Eq.~\eqref{eq:eft_bound} is not
satisfied unless $\varepsilon \approx 0.9$ (with only $j=0$ corrections) and
$\varepsilon \approx 0.7$ (with both $j=0$ and $1$ corrections).
%
The situation is different if we use $\gm = M_f$. In this case, we find that
with or without spin corrections that Eq.~\eqref{eq:eft_bound} can be satisfied
with $\varepsilon \leq 1/2$, i.e., below the criteria used
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}.
%
This means that with our model's assumptions and using the remnant's source mass $M_f$ to set the
cut-off scale that we can claim an upper bound
%
\begin{equation}
\ell_{\rm CS} \leqslant 46.5~\textrm{km}~( = \Lambda_{\rm EFT}(\tfrac{1}{2}, M_{f}) )
\quad \textrm{at 90\% credibility},
\end{equation}
%
on dCS gravity and would constitute the first bound on this theory with
gravitational-wave observations alone.
%
We also note that since the CDF grows fast in approximate domain $\varepsilon = [0.2, 0.4]$,
that a stronger, albeit at lower credibility,
%
\begin{equation}
\ell_{\rm CS} \leqslant 27.4~\textrm{km}~( = \Lambda_{\rm EFT}(0.3, M_{f}) )
\quad \textrm{at 68\% credibility},
\end{equation}
%
can also be placed, when we include the $j=1$ QNM correction.

We can draw qualitatively similar conclusions from the GW200129 event.
In particular, we find,
%
\begin{equation}
\ell_{\rm CS} \leqslant 43.9~\textrm{km}~( = \Lambda_{\rm EFT}(\tfrac{1}{2}, M_{f}) )
\quad \textrm{at 90\% credibility},
\end{equation}
%
and
%
\begin{equation}
\ell_{\rm CS} \leqslant 19.3~\textrm{km}~( = \Lambda_{\rm EFT}(0.3, M_{f}) )
\quad \textrm{at 68\% credibility},
\end{equation}
%
when we include the $j=1$ QNM correction.

These slightly stronger bound are a consequence of the larger support for $\ell_{\rm CS} \lessapprox 15$~km
for GW200129 (compare the top and middle panels in Fig.~\ref{fig:dCS_exec_sum}), and in part due to
the smaller median remmant ($M_{f} \approx 59.5\msun$ versus $M_{f} \approx 61.8\msun$ for GW150914).

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/dcs_cdf_varying_threshold.pdf}
\caption{(Color online).~The cumulative distribution function evaluated at the cut-off $\Lambda_{\rm EFT}(\varepsilon, \gm)$
as a function of the parameter $\varepsilon$ for both (blue curves) $m = m_2$, the secondary's source mass,
and (orange curves) $m = M_{f}$, the remnant's source mass, without (solid curve) and with (dashed curves)
linear in spin QNM corrections. The horizontal lines lay at 0.68 and 0.90.
%
We see that the CDF at $\varepsilon = 1/2$ is approximately one when $\gm =
M_{f}$ for both events, with and without including the $j = 1$ dCS corrections
to the dominat QNM.
}
\label{fig:dcs_cdf}
\end{figure}
